
---
title: 'Coding Session 1: Coding, Stats, and R Basics'
author: "Los Angeles City College -- Southern California Consortium for Data Science"
date: "Saturday November 2 2024"
output:
  pdf_document: default
  html_document:
    df_print: kable
---

\renewcommand{\P}{\mathds{P}}

```{r global_options, include=FALSE}
knitr::opts_chunk$set(out.width="70%", out.height="60%", fig.align="center", warning=FALSE, message=FALSE)
```

# 1. Basics

In this section, you will learn the basics of `R`.

To follow along, you should type the code that appears in each box to ensure you are getting the intended result.

1. When you encounter a "**Challenge**", this is a task you should solve with your group using critical thinking.
2. In this first part, whenever you see a box of code, you should also type out that code and run it to verify the output. We all need to learn the basics.
3. When we get to part 2 (*Data Frames*), it is okay to focus more on the challenges and read code and the output. However, you should still type out any code that feels "brand new." Physically typing out the code helps more than you might think.

Don't be afraid to ask for help at any point during the datathon! Raise your hand and a volunteer will come help right away.

## 1.1 `R` is a calculator

What is $2+2$? `R` can do this for us in three different ways.

### Method 1

Input `2+2` into the **console** and press `ENTER` to obtain `4`.

```{r}
2+2
```

### Method 2

Try doing it from the script by following these steps:

1. Input `2+2` into the **script** file.
2. Press `CTRL+ENTER`. *( `CMD+ENTER` on a Mac.)*

Now, look at the console and you will notice that the code "gets run" in the console.

### Method 3

* Go back to the line for `2+2` in the script file, and click the `RUN` button.

### Using Comments

Lines that start with a # are called *comments* and are not run. They are useful for making notes to yourself or to explain complex code.

Type each of these two lines into the script file and run them one-by-one.

```{r}
# Multiply 2 by 3 and then add 7
2*3 + 7
```

You do not need to type out the comments in this next example, but you should read them and run the code.

```{r}
# Calculate the average of four numbers.
# First, we sum the numbers using parentheses, then divide by the count.
(2 + 7 + 3 + 2) / 4
```

In what follows, you do not necessarily need to type out all of the comments. However, it is considered "good coding practice" to add comments to your code. *(Can you think of a couple of reasons as to why?)*
